---
title: "Exploratory Data Analysis"
format: html
editor: visual
---

(Optional) Package installations

```{r}
#| label: package install
#| warning: false
#| include: false

install.packages("tidyr")
install.packages("tidyverse")
install.packages("ggsoccer")
```

# Preprocessing

-   Unexpand

-   Removing NAs

```{r}
library(tidyr)
library(tidyverse)
library(ggsoccer)
library(knitr)
library(kableExtra)
library(ggplot2)
library(ggrepel)

# Read in full data set
pass360full <- readRDS("pass360full.rds")

# Passes that are from open play (no pressure in set pieces)
open_play_passes <- pass360full %>% 
  filter(play_pattern.name == "Regular Play" | play_pattern.name == "Other" | play_pattern.name == "From Counter" | play_pattern.name == "From Keeper")

# Data set with single open play passes (where the passer is the actor)
openplay_single =  open_play_passes %>% 
  filter(actor == TRUE, .keep_all = TRUE)

# Change NA pass outcome to complete
open_play_passes <- open_play_passes %>%
  mutate(pass.outcome.name = replace_na(pass.outcome.name, "Complete"))

openplay_single <- openplay_single %>%
  mutate(pass.outcome.name = replace_na(pass.outcome.name, "Complete"))

####################
# REMOVE NA Columns
###################

#Cannot remove player ID because that is the only true identifier for a player
cols_to_remove <- c("type.name", "type.id", "tactics.lineup", "under_pressure", "out",
                    "off_camera", "counterpress", "pass.switch", "pass.cross",
                    "pass.assisted_shot_id", "pass.shot_assist", "pass.aerial_won",
                    "pass.through_ball", "pass.outswinging", "pass.deflected",
                    "pass.no_touch", "pass.inswinging", "pass.goal_assist",
                    "pass.technique.id", "pass.technique.name", "carry.end_location",
                    "shot.end_location", "shot.freeze_frame", "goalkeeper.end_location",
                    "pass.cut_back", "pass.straight", "pass.miscommunication", "pass.recipient.name")

openplay_single <- openplay_single %>%
  select(-any_of(cols_to_remove))
open_play_passes <- open_play_passes %>%
  select(-any_of(cols_to_remove))

# Unexpand passes
pass_single =  pass360full %>% distinct(id, .keep_all = TRUE)

#Fix Bundesliga 
pass_single <- pass_single %>%
  mutate(
    competition_name = if_else(
      competition_name == "1. Bundesliga",
      "Bundesliga",
      competition_name
    )
  )

#Fix Bundesliga 
openplay_single <- openplay_single %>%
  mutate(
    competition_name = if_else(
      competition_name == "1. Bundesliga",
      "Bundesliga",
      competition_name
    )
  )
```

# Feature Engineering

### Condense into 4-3-3 formation

-   Center back will be assigned dynamically depending where they spend the majority of their time

-   Centre Midfield and Centre Attacking Midfield will be assigned the same way

```{r}
#| label: New Positions
#######################################
# Condense singles into 11 positions 
#######################################
openplay_single <- openplay_single %>%
  group_by(player.id) %>%
  mutate(
    avg_y = mean(players_y[position.name %in% c("Center Back", "Center Midfield", "Center Attacking Midfield")], na.rm = TRUE)
  ) %>%
  ungroup() %>%
  
  mutate(position_group = case_when(
    position.name == "Goalkeeper" ~ "Goalkeeper",

    position.name == "Left Center Back" ~ "Left Center Back",
    position.name == "Right Center Back" ~ "Right Center Back",
    
    position.name == "Center Back" ~ 
      if_else(!is.na(avg_y) & avg_y < 40, "Left Center Back", "Right Center Back"),
    
    position.name %in% c("Left Back", "Left Wing Back") ~ "Left Back",
    position.name %in% c("Right Back", "Right Wing Back") ~ "Right Back",
    
    position.name %in% c(
      "Left Defensive Midfield",
      "Center Defensive Midfield",
      "Right Defensive Midfield"
    ) ~ "Defensive Midfield",
    
    position.name %in% c(
      "Right Center Midfield",
      "Right Midfield",
      "Right Attacking Midfield"
    ) ~ "Right Central Midfield",
    
    position.name %in% c(
      "Left Center Midfield",
      "Left Midfield",
      "Left Attacking Midfield"
    ) ~ "Left Central Midfield",
    
    position.name %in% c("Center Midfield", "Center Attacking Midfield") ~ 
      if_else(!is.na(avg_y) & avg_y < 40, "Left Central Midfield", "Right Central Midfield"),
    
    position.name == "Left Wing" ~ "Left Wing",
    position.name == "Right Wing" ~ "Right Wing",
    
    position.name %in% c(
      "Left Center Forward",
      "Center Forward",
      "Right Center Forward"
    ) ~ "Striker",
    
    TRUE ~ position.name
  )) %>%
  select(-avg_y)


###################################################
# Condense freeze frame positions into 11 positions 
##################################################
open_play_passes <- open_play_passes %>%
  group_by(player.id) %>%
  mutate(
    avg_y = mean(players_y[position.name %in% c("Center Back", "Center Midfield", "Center Attacking Midfield")], na.rm = TRUE)
  ) %>%
  ungroup() %>%
  
  mutate(position_group = case_when(
    position.name == "Goalkeeper" ~ "Goalkeeper",
    
    position.name == "Left Center Back" ~ "Left Center Back",
    position.name == "Right Center Back" ~ "Right Center Back",
    
    position.name == "Center Back" ~ 
      if_else(!is.na(avg_y) & avg_y < 40, "Left Center Back", "Right Center Back"),
    
    position.name %in% c("Left Back", "Left Wing Back") ~ "Left Back",
    position.name %in% c("Right Back", "Right Wing Back") ~ "Right Back",
    
    position.name %in% c(
      "Left Defensive Midfield",
      "Center Defensive Midfield",
      "Right Defensive Midfield"
    ) ~ "Defensive Midfield",
    
    position.name %in% c(
      "Right Center Midfield",
      "Right Midfield",
      "Right Attacking Midfield"
    ) ~ "Right Central Midfield",
    
    position.name %in% c(
      "Left Center Midfield",
      "Left Midfield",
      "Left Attacking Midfield"
    ) ~ "Left Central Midfield",
    
    position.name %in% c("Center Midfield", "Center Attacking Midfield") ~ 
      if_else(!is.na(avg_y) & avg_y < 40, "Left Central Midfield", "Right Central Midfield"),
    
    position.name == "Left Wing" ~ "Left Wing",
    position.name == "Right Wing" ~ "Right Wing",
    
    position.name %in% c(
      "Left Center Forward",
      "Center Forward",
      "Right Center Forward"
    ) ~ "Striker",
    
    TRUE ~ position.name
  )) %>%
  select(-avg_y)
```

```{r}
#| label: In frame
#| warning: false
########################################################
### To check if the end location is in the visbile area
#################################################
library(sp) 
library(purrr)

is_in_visible_area <- function(visible_area, end_x, end_y) {
  if (is.null(visible_area) || length(visible_area) < 6 || is.na(end_x) || is.na(end_y)) {
    return(NA)
  }
  
  # split flat vector into x's (odd indices) and y's (even indices)
  xs <- visible_area[seq(1, length(visible_area), by = 2)]
  ys <- visible_area[seq(2, length(visible_area), by = 2)]
  
  result <- point.in.polygon(end_x, end_y, xs, ys)
  result > 0  # 0 = outside, 1 = inside, 2/3 = on edge/vertex
}

openplay_single <- openplay_single %>%
  mutate(
    end_in_visible_area = pmap_lgl(
      list(visible_area, pass_end_x, pass_end_y),
      is_in_visible_area
    )
  )
open_play_passes <- open_play_passes %>%
  mutate(
    end_in_visible_area = pmap_lgl(
      list(visible_area, pass_end_x, pass_end_y),
      is_in_visible_area
    )
  )
```

```{r}
#| label: Pass distance, angle, transition, gametime 

openplay_single = openplay_single %>%
  # Categorical distance (short, medium, long)
  mutate(distance = case_when(
    pass.length < 15 ~ "Short",
    pass.length >= 15 & pass.length < 30 ~ "Medium",
    pass.length >= 30 ~ "Long",
    TRUE ~ NA_character_
  ),
  # Categorical direction (Backwards, Sideways, Forwards)
  direction = case_when(
    abs(pass.angle) <= pi/2 ~ "Forwards",
    abs(pass.angle) > 0.75*pi ~ "Backwards",
    TRUE ~ "Sideways"
  ),
  # Categorical Gametime (1st, 2nd, 3rd, 4th, ET1, ET2)
  quarter = case_when(
    period == 1 & minute < 23  ~ "Q1",
    period == 1 & minute >= 23 ~ "Q2",
    period == 2 & minute < 68  ~ "Q3",   # second half starts ~45, so midpoint ~67-68
    period == 2 & minute >= 68 ~ "Q4",
    period == 3               ~ "ET1",   # first extra-time half
    period == 4               ~ "ET2",   # second extra-time half
    TRUE ~ NA_character_
  ),
  # Categorical space transition (Defence, Mid, Attack)
  transition = case_when(
    players_x < 40 & pass_end_x < 40 ~ "Defence - Defence",
    players_x < 40 & pass_end_x >= 40 & pass_end_x < 80 ~ "Defence - Mid",
    players_x < 40 & pass_end_x >= 80 ~ "Defence - Attack",          # long ball forward
    players_x >= 40 & players_x < 80 & pass_end_x < 40 ~ "Mid - Defence",   # back pass
    players_x >= 40 & players_x < 80 & pass_end_x >= 40 & pass_end_x < 80 ~ "Mid - Mid",
    players_x >= 40 & players_x < 80 & pass_end_x >= 80 ~ "Mid - Attack",
    players_x >= 80 & pass_end_x < 40 ~ "Attack - Defence",          # long ball back
    players_x >= 80 & pass_end_x >= 40 & pass_end_x < 80 ~ "Attack - Mid",  # back pass
    players_x >= 80 & pass_end_x >= 80 ~ "Attack - Attack",
    TRUE ~ NA_character_
  )
  )
```

## Voronoi Areas

```{r}

# Group by id 
# Compute Vor object for pass event using deldir
# Obtain tile list using tile.list
# 

# Join in openplay_single using player_id
tile.list(deldir((open_play_passes$players_x)[1:16], (open_play_passes$players_y)[1:16], rw = c(0, 120, 0, 80)))$pt.3$area
```

# Exploratory Data Analysis

-   Number of passes, percentage of successful passes, competition, season

## Position Analysis

```{r}
position_counts = open_play_passes %>%
  group_by(Position = position_group) %>%
  summarise(
    Number_of_players = n_distinct(player.name),
    .groups = "drop"
  ) %>%
  arrange(desc(Number_of_players))

position_table = kable(
  position_counts,
  format = "latex",
  booktabs = TRUE,
  caption = "Number of players in each position",
  label = "position_counts"
)%>%
  kable_styling(
    latex_options = "hold_position"
  )

cat(position_table)
```

## Pass completion analysis

### Open Play

Analysing pass completion segmented by:

Distance: Short (\<15m), medium (\<30m), long (\>30m)

Angle: Forwards (0-45), sideways (45-135), backwards (135-180)

Height: Ground, Low, High

Sector: Defence, Mid, Final third

Game context: home/away, losing/leading (don't have this in pass data), can do game time in quarters

```{r}
#| label: Passing accuracy
#| warning: false

summary_table = openplay_single %>% 
  mutate(distance = case_when(
    pass.length < 15 ~ "Short",
    pass.length >= 15 & pass.length < 30 ~ "Medium",
    pass.length >= 30 ~ "Long",
    TRUE ~ NA_character_
  ),
  angle = case_when(
    abs(pass.angle) <= pi/2 ~ "Forwards",
    abs(pass.angle) > 0.75*pi ~ "Backwards",
    TRUE ~ "Sideways"
  ),
  sector = case_when(
    players_x < 40 & pass_end_x < 40 ~ "Defence-Defence",
    players_x < 40 & (pass_end_x >= 40 & pass_end_x < 80) ~ "Defence - Mid",
    players_x >= 40 & players_x < 80 & pass_end_x >= 40 & pass_end_x < 80 ~ "Mid - Mid",
    players_x >= 40 & players_x < 80 & pass_end_x >= 80 ~ "Mid - Attack",
    TRUE ~ "Attack - Attack"
  )
  )

make_summary <- function(data, group_var) {
  data %>%
    group_by(category = {{ group_var }}) %>%
    summarise(
      Total = n(),
      Successful = sum(pass.outcome.name == "Complete", na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(
      Proportion = round(100 * Total / nrow(data), 2),
      Percentage = round(100 * Successful / Total, 2)
    ) %>%
    relocate(Proportion, .after = Total)
}

df <- summary_table %>%
  summarise(category = "All Passes", Total = n(),
            Successful = sum(pass.outcome.name == "Complete", na.rm = TRUE)) %>%
  mutate(Proportion = round(100 * Total / nrow(summary_table), 2),
         Percentage = round(100 * Successful / Total, 2)) %>% 
  relocate(Proportion, .after = Total)


distance_tbl <- make_summary(summary_table, distance) %>% 
  arrange(desc(Percentage))
direction_tbl    <- make_summary(summary_table, angle) %>% 
  arrange(desc(Percentage))
transition_tbl   <- make_summary(summary_table, sector) %>% 
  arrange(desc(Percentage))
height_tbl <- make_summary(summary_table, pass.height.name) %>% 
  arrange(desc(Percentage))

final_summary_table <- bind_rows(df, direction_tbl, distance_tbl,height_tbl, transition_tbl)

printtable = kable(
  final_summary_table,
  format = "latex",
  booktabs = TRUE,
  caption = "Passing Accuracy",
  label = "position_counts"
)
cat(printtable)

quarter_tbl <- make_summary(summary_table, quarter)
quarter_table <- bind_rows(df, quarter_tbl)
printquarter = kable(
  quarter_table,
  format = "latex",
  booktabs = TRUE,
  caption = "Passing Accuracy per Quarter",
  label = "position_counts"
)
cat(printquarter)
```

```{r}
#| label: rankings

###########################################
# Table of passing frequencies and accuracy
############################################
tot_pass_accuracies <- openplay_single %>%
  group_by(player.id, position_group) %>%
  summarise(
    Total = n(),
    Successful = sum(pass.outcome.name == "Complete", na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(Accuracy = round(100 * Successful / Total, 2)) %>% 
  select(-Successful)

################################
# Top player computations
#################################
top5_frequency <- tot_pass_accuracies %>%
  arrange(desc(Total)) %>%
  slice_head(n = 5)

top5_frequency_kable = kable(
  top5_frequency,
  format = "latex",
  booktabs = TRUE,
  caption = "Top 5 Overall Passing Frequency",
  label = "top5Overall_pass_freq"
)
cat(top5_frequency_kable)

top5_accuracy_min300 <- tot_pass_accuracies %>%
  filter(Total >= 286) %>%
  arrange(desc(Accuracy)) %>%
  slice_head(n = 5)
top5_accuracy_kable = kable(
  top5_accuracy_min300,
  format = "latex",
  booktabs = TRUE,
  caption = "Top 5 Overall Passing Accuracy",
  label = "top5Overall_pass_acc"
)
cat(top5_accuracy_kable)

top5_frequency
# Xhaka, Tah, Wirtz, Palacios, Tapsoba
top5_accuracy_min300
# Christensen, Tah, Rodri, Laporte, Ruben Dias

#####################
# Per position plots
#####################
# top3_position <- tot_pass_accuracies %>%
#   filter(Total >= 140, position_group != "Goalkeeper") %>%       
#   group_by(position_group) %>%
#   slice_max(Accuracy, n = 3, with_ties = FALSE) %>%
#   ungroup() %>% 
#   arrange(desc(Accuracy)) %>%
#   mutate(rank = row_number()) # unique y position per player, ordered by Accuracy
# 
# ggplot(top3_position,
#        aes(x = Total, y = Accuracy, colour = position_group)) +
#   geom_point(size = 2) +
#   geom_text_repel(
#     aes(label = player.id),
#     size = 3,
#     show.legend = FALSE,
#     box.padding = 0.3,
#     max.overlaps = Inf
#   ) +
#   scale_colour_brewer(palette = "Dark2") +
#   labs(
#     x = "Total Passes",
#     y = "Passing Accuracy (%)",
#     colour = "Position",
#     title = "Passing Accuracy vs. Volume by Position"
#   ) +
#   theme_minimal(base_size = 12) +
#   theme(
#     panel.grid.minor = element_blank()
#   )

top10_position <- tot_pass_accuracies %>%
  filter(Total >= 140, position_group != "Goalkeeper", player.id != 3500) %>%       
  group_by(position_group) %>%
  slice_max(Accuracy, n = 10, with_ties = FALSE) %>%
  ungroup() %>% 
  arrange(desc(Accuracy)) %>%
  mutate(rank = row_number()) # unique y position per player, ordered by Accuracy

ggplot(top10_position,
       aes(x = Total, y = Accuracy, colour = position_group)) +
  geom_point(size = 2) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    x = "Total Passes",
    y = "Passing Accuracy (%)",
    colour = "Position",
    title = "Passing Accuracy vs. Volume by Position"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank()
  )
```

### All Passes or Set Pieces

```{r}

```

## Competition Analysis

### Competition Passes Table (All Passes)

-   Passes along with the competiton, season name, number of passes, average number of passes per match

```{r}
pass_summary = openplay_single %>% group_by(Competition = competition_name,Season = season_name,) %>% 
  summarise(
    Total = n(),
    Matches = n_distinct(match_id),
    Passes_per_match = round(Total / Matches, 2),
    .groups = "drop"
  ) %>%
  arrange(desc(Passes_per_match))

pass_summary

pass_table <- kable(
  pass_summary,
  format = "latex",
  booktabs = TRUE,
  caption = "Number of passes by season and competition (Open)",
  label = "tab:pass_summary",
  longtable = TRUE
) %>%
  kable_styling(
    latex_options = "hold_position"
  )

cat(pass_table)
```

### Open Play

```{r}
pass_summary = openplay_single %>% group_by(Competition = competition_name,Season = season_name,) %>% 
  summarise(
     Total = n(),
    Matches = n_distinct(match_id),
    Passes_per_match = round(Total / Matches, 2),
    Successful = sum(pass.outcome.name == "Complete", na.rm = TRUE),
    .groups = "drop"
  )  |>  mutate(Percentage = round(100 * Successful / Total, 2)) |> 
  arrange(desc(Passes_per_match))

pass_summary

pass_table <- kable(
  pass_summary,
  format = "latex",
  booktabs = TRUE,
  caption = "Number of passes by season and competition(Open Play)",
  label = "tab:pass_summary",
  longtable = TRUE
)

cat(pass_table)
```

### Seperate into positions

#### Pass Completion by position

```{r}
position_summary = openplay_single %>% group_by(Position = position_group) %>% 
  summarise(
    Total = n(),
    Matches = n_distinct(match_id),
    Passes_per_match = round(Total / Matches, 2),
    Successful = sum(pass.outcome.name == "Complete", na.rm = TRUE),
    .groups = "drop"
  )  |>  mutate(Percentage = round(100 * Successful / Total, 2)) |> 
  arrange(desc(Passes_per_match))

position_summary

position_table <- kable(
  head(position_summary,11),
  format = "latex",
  booktabs = TRUE,
  caption = "Number of passes by position",
  label = "tab:position_summary",
  longtable = TRUE
)

cat(position_table)
```

### Separating Passes into competitions

#### Bundesliga

```{r}
#Bundesliga 2023/2024
bund_2324 = subset(openplay_single, competition_name == "Bundesliga")

bund_summary <- bund_2324 %>%
  group_by(
    Position = position.name
  ) %>%
  summarise(
    Total = n(),
    Matches = n_distinct(match_id),
    Passes_per_match = round(Total / Matches, 2),
    .groups = "drop"
  ) %>%
  arrange(desc(Total))

bund_table <- kable(
  bund_summary,
  format = "latex",
  booktabs = TRUE,
  caption = "Number of passes by position in the Bundesliga 2023/24",
  label = "tab:bund_table",
  longtable = TRUE
)

cat(bund_table)
```

#### Fifa World Cup 2022

```{r}
#Fifa World Cup 2022
wc_2022 = subset(openplay_single, competition_name == "FIFA World Cup")

wc_2022_summary <- wc_2022 %>%
  group_by(
    PlayerID = player.id,
    Position = position.name
  ) %>%
  summarise(
    Total = n(),
    Matches = n_distinct(match_id),
    Passes_per_match = round(Total / Matches, 2),
    .groups = "drop"
  ) %>%
  arrange(desc(Total))

wc_2022_table <- kable(
  bund_summary,
  format = "latex",
  booktabs = TRUE,
  caption = "Number of passes by position in the Fifa World Cup 2022",
  label = "tab:wc_2022_table",
  longtable = TRUE
)

writeLines(wc_2022_table, "wc_2022_table.tex")
```

```{r}
#UEFA Euro 2020
euro_2020 = subset(pass_single,competition_name == "UEFA Euro" & season_name=="2020")

#UEFA Euro 2024
euro_2024 = subset(pass_single,competition_name == "UEFA Euro" & season_name=="2024")

#UEFA Women's Euro 2022
euro_w_2022 = subset(pass_single,competition_name == "UEFA Women's Euro" & season_name=="2022")

#UEFA Women's Euro 2025
euro_w_2025 = subset(pass_single,competition_name == "UEFA Women's Euro" & season_name=="2025")

#Womans World Cup 2023
wc_w_2023 = subset(pass_single,competition_name == "Women's World Cup")
```

## Heatmaps

-   Segmented by competition and season

### Heatmap of pass completion

```{r}

```

### Heatmaps segmented by position

#### Centre Backs (Open Play Passes)

```{r}
centre_back_data = openplay_single %>% filter(grepl("Center Back", position.name))

pass_data = data.frame(
  x = centre_back_data$players_x,
  y =  centre_back_data$players_y
)


ggplot(pass_data, aes(x = x, y = y)) +
  annotate_pitch(
    dimensions = pitch_statsbomb,
    colour = "white",
    fill = "springgreen4",
    limits = F,
    alpha = 1
  ) +
  geom_bin2d(
    binwidth = c(5, 5),
    alpha = 0.85
  ) +
  scale_fill_gradient(
    low = "yellow", 
    high = "red",
    name = "Pass Count"
  ) +
  coord_cartesian(xlim = c(0, 120), ylim = c(0, 80)) +
  theme_pitch() +
  labs(
    title = "Pass Heatmap for Centre Backs",
    subtitle = "Segmented into 5 by 5 metre squares"
  ) 

range(pass_data$x)
range(pass_data$y)
```

#### Centre Backs Heatmap

```{r}
ggplot(pass_data, aes(x = x, y = y)) +
  annotate_pitch(
    dimensions = pitch_statsbomb,
    colour = "white",
    fill = "springgreen4",
    limits = F,
    alpha = 1
  ) +
  stat_density2d(
    aes(fill = ..level..), 
    geom = "polygon", 
    alpha = 0.3, 
    bins = 20,
  ) +
  coord_cartesian(xlim = c(0, 120), ylim = c(0, 80))+
  scale_fill_gradient(low = "yellow", high = "red",name = "Pass Density") +
  theme_pitch() +
  labs(title = "Pass Heatmap Centre Backs (Open Play)") +
  scale_x_continuous(
  limits = c(0, 120),
  breaks = seq(0, 120, 20)
) +
scale_y_continuous(
  limits = c(0, 80),
  breaks = seq(0, 80, 10)
)
```

#### Wing Backs Grid

```{r}
wing_back_data = openplay_single %>% filter(grepl("Wing Back", position.name))

pass_data = data.frame(
  x = wing_back_data$players_x,
  y =  wing_back_data$players_y
)


ggplot(pass_data, aes(x = x, y = y)) +
  annotate_pitch(
    dimensions = pitch_statsbomb,
    colour = "white",
    fill = "springgreen4",
    limits = F,
    alpha = 1
  ) +
  geom_bin2d(
    binwidth = c(5, 5),
    alpha = 0.85
  ) +
  scale_fill_gradient(
    low = "yellow", 
    high = "red",
    name = "Pass Count"
  ) +
  coord_cartesian(xlim = c(0, 120), ylim = c(0, 80)) +
  theme_pitch() +
  labs(
    title = "Pass Heatmap for Wing Backs",
    subtitle = "Segmented into 5 by 5 metre squares"
  ) 
```

#### Wingback Heatmap

```{r}
ggplot(pass_data, aes(x = x, y = y)) +
  annotate_pitch(
    dimensions = pitch_statsbomb,
    colour = "white",
    fill = "springgreen4",
    limits = F,
    alpha = 1
  ) +
  stat_density2d(
    aes(fill = ..level..), 
    geom = "polygon", 
    alpha = 0.3, 
    bins = 20,
  ) +
  coord_cartesian(xlim = c(0, 120), ylim = c(0, 80))+
  scale_fill_gradient(low = "yellow", high = "red",name = "Pass Density") +
  theme_pitch() +
  labs(title = "Pass Heatmap Left and right wing backs (Open Play)") +
  scale_x_continuous(
  limits = c(0, 120),
  breaks = seq(0, 120, 20)
) +
scale_y_continuous(
  limits = c(0, 80),
  breaks = seq(0, 80, 10)
)
```

### Left/Right Backs (Grid)

```{r}
back_data = openplay_single %>% filter(position.name == "Left Back" |position.name == "Right Back") 

pass_data = data.frame(
  x = back_data$players_x,
  y =  back_data$players_y
)


ggplot(pass_data, aes(x = x, y = y)) +
 annotate_pitch(
    dimensions = pitch_statsbomb,
    colour = "white",
    fill = "springgreen4",
    limits = F,
    alpha = 1
  ) +
  geom_bin2d(
    binwidth = c(5, 5),
    alpha = 0.85
  ) +
  scale_fill_gradient(
    low = "yellow", 
    high = "red",
    name = "Pass Count"
  ) +
  coord_cartesian(xlim = c(0, 120), ylim = c(0, 80)) +
  theme_pitch() +
  labs(
    title = "Pass Heatmap for Left and Right Backs",
    subtitle = "Segmented into 5 by 5 metre squares"
  ) 
```

#### Left/Right Backs Heatmap

```{r}
ggplot(pass_data, aes(x = x, y = y)) +
  annotate_pitch(
    dimensions = pitch_statsbomb,
    colour = "white",
    fill = "springgreen4",
    limits = F,
    alpha = 1
  ) +
  stat_density2d(
    aes(fill = ..level..), 
    geom = "polygon", 
    alpha = 0.3, 
    bins = 20,
  ) +
  coord_cartesian(xlim = c(0, 120), ylim = c(0, 80))+
  scale_fill_gradient(low = "yellow", high = "red",name = "Pass Density") +
  theme_pitch() +
  labs(title = "Pass Heatmap Left and Right Backs (Open Play)") +
  scale_x_continuous(
  limits = c(0, 120),
  breaks = seq(0, 120, 20)
) +
scale_y_continuous(
  limits = c(0, 80),
  breaks = seq(0, 80, 10)
)
```

#### Midfielders (Grid)

```{r}
midfield_data = openplay_single %>% filter(grepl("Midfield", position.name))

pass_data = data.frame(
  x =midfield_data$players_x,
  y =midfield_data$players_y
)


ggplot(pass_data, aes(x = x, y = y)) +
  annotate_pitch(
    dimensions = pitch_statsbomb,
    colour = "white",
    fill = "springgreen4",
    limits = F,
    alpha = 1
  ) +
  geom_bin2d(
    binwidth = c(5, 5),
    alpha = 0.85
  ) +
  scale_fill_gradient(
    low = "yellow", 
    high = "red",
    name = "Pass Count"
  ) +
  coord_cartesian(xlim = c(0, 120), ylim = c(0, 80)) +
  theme_pitch() +
  labs(
    title = "Pass Heatmap for Midfielders",
    subtitle = "Segmented into 5 by 5 metre squares"
  ) 
```

#### Midfielders (Heatmap)

```{r}
ggplot(pass_data, aes(x = x, y = y)) +
  annotate_pitch(
    dimensions = pitch_statsbomb,
    colour = "white",
    fill = "springgreen4",
    limits = F,
    alpha = 1
  ) +
  stat_density2d(
    aes(fill = ..level..), 
    geom = "polygon", 
    alpha = 0.3, 
    bins = 20,
  ) +
  coord_cartesian(xlim = c(0, 120), ylim = c(0, 80))+
  scale_fill_gradient(low = "yellow", high = "red",name = "Pass Density") +
  theme_pitch() +
  labs(title = "Pass Heatmap Midfielders (Open Play)") +
  scale_x_continuous(
  limits = c(0, 120),
  breaks = seq(0, 120, 20)
) +
scale_y_continuous(
  limits = c(0, 80),
  breaks = seq(0, 80, 10)
)
```

#### Wingers (Grid)

```{r}
wing_data = openplay_single %>% filter(position.name == "Left Wing" |position.name == "Right Wing") 

pass_data = data.frame(
  x = wing_data$players_x,
  y =  wing_data$players_y
)


ggplot(pass_data, aes(x = x, y = y)) +
  annotate_pitch(
    dimensions = pitch_statsbomb,
    colour = "white",
    fill = "springgreen4",
    limits = F,
    alpha = 1
  ) +
  geom_bin2d(
    binwidth = c(5, 5),
    alpha = 0.85
  ) +
  scale_fill_gradient(
    low = "yellow", 
    high = "red",
    name = "Pass Count"
  ) +
  coord_cartesian(xlim = c(0, 120), ylim = c(0, 80)) +
  theme_pitch() +
  labs(
    title = "Pass Heatmap for Left and Right Wings",
    subtitle = "Segmented into 5 by 5 metre squares"
  ) 
```

#### Wingers (Heatmap)

```{r}
ggplot(pass_data, aes(x = x, y = y)) +
  annotate_pitch(
    dimensions = pitch_statsbomb,
    colour = "white",
    fill = "springgreen4",
    limits = F,
    alpha = 1
  ) +
  stat_density2d(
    aes(fill = ..level..), 
    geom = "polygon", 
    alpha = 0.3, 
    bins = 20,
  ) +
  coord_cartesian(xlim = c(0, 120), ylim = c(0, 80))+
  scale_fill_gradient(low = "yellow", high = "red",name = "Pass Density") +
  theme_pitch() +
  labs(title = "Pass Heatmap Wingers (Open Play)") +
  scale_x_continuous(
  limits = c(0, 120),
  breaks = seq(0, 120, 20)
) +
scale_y_continuous(
  limits = c(0, 80),
  breaks = seq(0, 80, 10)
)
```

#### Forwards (Grid)

```{r}
forward_data = openplay_single %>% filter(grepl("Forward", position.name))

pass_data = data.frame(
  x =forward_data$players_x,
  y =forward_data$players_y
)


ggplot(pass_data, aes(x = x, y = y)) +
  annotate_pitch(
    dimensions = pitch_statsbomb,
    colour = "white",
    fill = "springgreen4",
    limits = F,
    alpha = 1
  ) +
  geom_bin2d(
    binwidth = c(5, 5),
    alpha = 0.85
  ) +
  scale_fill_gradient(
    low = "yellow", 
    high = "red",
    name = "Pass Count"
  ) +
  coord_cartesian(xlim = c(0, 120), ylim = c(0, 80)) +
  theme_pitch() +
  labs(
    title = "Pass Heatmap for Forwards",
    subtitle = "Segmented into 5 by 5 metre squares"
  ) 
```

#### Forwards (Heatmap)

```{r}
ggplot(pass_data, aes(x = x, y = y)) +
  annotate_pitch(
    dimensions = pitch_statsbomb,
    colour = "white",
    fill = "springgreen4",
    limits = F,
    alpha = 1
  ) +
  stat_density2d(
    aes(fill = ..level..), 
    geom = "polygon", 
    alpha = 0.3, 
    bins = 15,
  ) +
  coord_cartesian(xlim = c(0, 120), ylim = c(0, 80))+
  scale_fill_gradient(low = "yellow", high = "red",name = "Pass Density") +
  theme_pitch() +
  labs(title = "Pass Heatmap Strikers (Open Play)") +
  scale_x_continuous(
  limits = c(0, 120),
  breaks = seq(0, 120, 20)
) +
scale_y_continuous(
  limits = c(0, 80),
  breaks = seq(0, 80, 10)
)
```

#### Goalkeeper (Grid)

```{r}
goalkeeper_data = openplay_single %>% filter(grepl("Goalkeeper", position.name))

pass_data = data.frame(
  x =goalkeeper_data$players_x,
  y =goalkeeper_data$players_y
)


ggplot(pass_data, aes(x = x, y = y)) +
  annotate_pitch(
    dimensions = pitch_statsbomb,
    colour = "white",
    fill = "springgreen4",
    limits = F,
    alpha = 1
  ) +
  geom_bin2d(
    binwidth = c(5, 5),
    alpha = 0.85
  ) +
  scale_fill_gradient(
    low = "yellow", 
    high = "red",
    name = "Pass Count"
  ) +
  coord_cartesian(xlim = c(0, 120), ylim = c(0, 80)) +
  theme_pitch() +
  labs(
    title = "Pass Heatmap for Goalkeepers",
    subtitle = "Segmented into 5 by 5 metre squares"
  ) 
```

#### Goalkeeper Heatmap

```{r}
ggplot(pass_data, aes(x = x, y = y)) +
  annotate_pitch(
    dimensions = pitch_statsbomb,
    colour = "white",
    fill = "springgreen4",
    limits = F,
    alpha = 1
  ) +
  stat_density2d(
    aes(fill = ..level..), 
    geom = "polygon", 
    alpha = 0.3, 
    bins = 15,
  ) +
  coord_cartesian(xlim = c(0, 120), ylim = c(0, 80))+
  scale_fill_gradient(low = "yellow", high = "red",name = "Pass Density") +
  theme_pitch() +
  labs(title = "Pass Heatmap Goalkeepers (Open Play)") +
  scale_x_continuous(
  limits = c(0, 120),
  breaks = seq(0, 120, 20)
) +
scale_y_continuous(
  limits = c(0, 80),
  breaks = seq(0, 80, 10)
)
```

### Heatmap of pass frequency for all open play passes

```{r}
pass_data = data.frame(
  x =openplay_single$players_x,
  y =openplay_single$players_y
)


ggplot(pass_data, aes(x = x, y = y)) +
  annotate_pitch(
    dimensions = pitch_statsbomb,
    colour = "white",
    fill = "springgreen4",
    limits = F,
    alpha = 1
  ) +
  stat_density2d(
    aes(fill = ..level..), 
    geom = "polygon", 
    alpha = 0.3, 
    bins = 15,
  ) +
  coord_cartesian(xlim = c(0, 120), ylim = c(0, 80))+
  scale_fill_gradient(low = "yellow", high = "red",name = "Pass Density") +
  theme_pitch() +
  labs(title = "Pass Heatmap All Positions (Open Play)") +
  scale_x_continuous(
  limits = c(0, 120),
  breaks = seq(0, 120, 20)
) +
scale_y_continuous(
  limits = c(0, 80),
  breaks = seq(0, 80, 10)
)



```

### Heatmap as grid

```{r}
pass_data <- data.frame(
  x = openplay_single$players_x,
  y = openplay_single$players_y
)

ggplot(pass_data, aes(x = x, y = y)) +
  annotate_pitch(
    dimensions = pitch_statsbomb,
    colour = "white",
    fill = "springgreen4",
    limits = F,
    alpha = 1
  ) +
  geom_bin2d(
    binwidth = c(5, 5),
    alpha = 0.85
  ) +
  scale_fill_gradient(
    low = "yellow", 
    high = "red",
    name = "Pass Count"
  ) +
  coord_cartesian(xlim = c(0, 120), ylim = c(0, 80)) +
  theme_pitch() +
  labs(
    title = "Pass Heatmap (Grid)",
    subtitle = "Segmented into 5 by 5 metre squares"
  ) 
```

## Distributions

-   When a pass has NA as the outcome, it was completed

### Distribution of Pass Outcomes

-   Segmented into different games and different leagues

#### Distribution of Pass Outcomes Open Play

```{r}

pie_data <- openplay_single %>%
  mutate(outcome = ifelse(is.na(pass.outcome.name), "Complete", pass.outcome.name)) %>%
  count(outcome)

# 2. Build the Pie Chart
ggplot(pie_data, aes(x = "", y = n, fill = outcome)) +
  # Create a stacked bar chart first
  geom_bar(stat = "identity", width = 1, color = "white") +
  # Convert the bar chart into a pie chart circle
  coord_polar("y", start = 0) +
  # Clean up the theme (removes gridlines and axis ticks)
  theme_void() + 
  labs(
    title = "Pass Outcomes Distribution (Open Play)",
    fill = "Outcome"
  ) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.title = element_text(face = "bold")
  )


```

#### Distribution of Pass Outcomes All Passes

```{r}
pie_data <- pass_single %>%
  mutate(outcome = ifelse(is.na(pass.outcome.name), "Complete", pass.outcome.name)) %>%
  count(outcome)

# 2. Build the Pie Chart
ggplot(pie_data, aes(x = "", y = n, fill = outcome)) +
  # Create a stacked bar chart first
  geom_bar(stat = "identity", width = 1, color = "white") +
  # Convert the bar chart into a pie chart circle
  coord_polar("y", start = 0) +
  # Clean up the theme (removes gridlines and axis ticks)
  theme_void() + 
  labs(
    title = "Pass Outcomes Distribution",
    fill = "Outcome"
  ) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.title = element_text(face = "bold")
  )
```

```{r}
#| label: check-columns
#| echo: false

###############################################################
# Check amount of non-missing values in each variable (column)
###############################################################
props <- c()

for (col in 1:ncol(openplay_single)) {
  props[col] <- ((sum(!is.na(openplay_single[,col])))/(nrow(openplay_single)))*100
}

col_counts <- data.frame(colname = colnames(openplay_single), percentage = props)
col_counts[order(col_counts$percentage),]

# Keep columns with 95% of data available
cols_to_keep <- col_counts$colname[col_counts$percentage >= 95]
openplay <- open_play_passes %>% 
  select(all_of(cols_to_keep))

# Some columns still have NULL only values
cols_to_drop <- c("tactics.lineup","carry.end_location", "shot.end_location", "shot.freeze_frame", "goalkeeper.end_location")
openplay <- openplay %>% 
  select(-cols_to_drop)

###########################
# Investigate pass outcome
###########################

uniq_outcomes <- unique(openplay_single$pass.outcome.name)
outcome_count <- c()

for (type in 1:length(uniq_outcomes)) {
  outcome_count[type] <- pass_single %>% filter(pass.outcome.name == uniq_outcomes[type]) %>% summarise(n = n()) %>% pull(n)
}

outcome_counts <- data.frame(colname = uniq_outcomes, count = outcome_count)
outcome_counts$count[1] = sum(is.na(pass_single$pass.outcome.name))
# TABLE OF OUTCOME DISTRIBUTION
outcome_counts[order(outcome_counts$count),]
barplot(outcome_counts$count, names.arg = outcome_counts$colname)

#####################################
# Change NA pass outcome to complete
#####################################

open_play_passes <- open_play_passes %>%
 mutate(pass.outcome.name = replace_na(pass.outcome.name, "Complete"))

openplay_single <- openplay_single %>%
 mutate(pass.outcome.name = replace_na(pass.outcome.name, "Complete"))

##########################################
# Update pass outcome in total and singles
##########################################

uniq_outcomes <- unique(openplay_single$pass.outcome.name)
outcome_count <- c()

for (type in 1:length(uniq_outcomes)) {
  outcome_count[type] <- pass_single %>% filter(pass.outcome.name == uniq_outcomes[type]) %>% summarise(n = n()) %>% pull(n)
}

outcome_counts <- data.frame(colname = uniq_outcomes, count = outcome_count)
outcome_counts$count[1] = sum(is.na(pass_single$pass.outcome.name))
# TABLE OF OUTCOME DISTRIBUTION
outcome_counts[order(outcome_counts$count),]
barplot(outcome_counts$count, names.arg = outcome_counts$colname, main = "Distribution of Pass Outcomes", ylab = "Count")
```

### Distribution of Pass Angles

-   Segmented into different games and different league

#### Open Play

```{r}
hist(openplay_single$pass.angle, main = "Pass Angle Distribution", xlab = "Angle (radians)")

ggplot(openplay_single, aes(x = pass.angle, fill = interaction(competition_name, season_name))) +
  geom_histogram(
    binwidth = 0.2, 
    color = "white", 
    alpha = 0.85,
  ) +
  scale_x_continuous(
    breaks = c(-pi, -pi/2, 0, pi/2, pi),
    labels = c("-π", "-π/2", "0", "π/2", "π"),
    limits = c(-3.2, 3.2)
  ) +
 labs(
    title = "Pass Angle Distribution by Competition & Season (Open Play)",
    x = "Angle (radians)",
    y = "Count"
  )+
  facet_wrap(~ competition_name + season_name, ncol = 3) +
  theme(legend.position = "none")
```

#### All Passes

```{r}
library(gghighlight)
ggplot(pass_single, aes(x = pass.angle, fill = interaction(competition_name, season_name))) +
  geom_histogram(
    binwidth = 0.2, 
    color = "white", 
    alpha = 0.85
  ) +
  scale_x_continuous(
    breaks = c(-pi, -pi/2, 0, pi/2, pi),
    labels = c("-π", "-π/2", "0", "π/2", "π"),
    limits = c(-3.2, 3.2)
  ) + 
  scale_fill_discrete(
    name = "Competition and Season Name",                                
  ) +
 labs(
    title = "Pass Angle Distribution by Competition & Season",
    x = "Angle (radians)",
    y = "Count"
  )
```

### Distribution of Pass Lengths

-   Segmented into different games and different leagues

#### Open Play

```{r}

ggplot(openplay_single, aes(x = pass.length, fill = interaction(competition_name, season_name))) +
  geom_histogram(
    binwidth = 2, 
    color = "white", 
    alpha = 0.85
  ) +
  scale_fill_discrete(
    name = "Competition and Season Name",                                
  ) +
 labs(
    title = "Pass Length Distribution by Competition & Season (Open Play)",
    x = "Length (yards)",
    y = "Count"
  ) +
  facet_wrap(~ competition_name + season_name, ncol = 3) +
  theme(legend.position = "none")
```

#### All Passes

```{r}
ggplot(pass_single, aes(x = pass.length, fill = interaction(competition_name, season_name))) +
  geom_histogram(
    binwidth = 2, 
    color = "white", 
    alpha = 0.85
  ) +
  scale_fill_discrete(
    name = "Competition and Season Name",                                
  ) +
 labs(
    title = "Pass Length Distribution by Competition & Season",
    x = "Length (metres)",
    y = "Count"
  )
```

## Freeze Frames

## Line graphs

### Relationship between passes and goals scored

### Pass completion as function of length and angle

## Pass Completion vs Length

### Using all labels for pass outcome

```{r}
plot_data <- pass_single %>%
  mutate(outcome = ifelse(is.na(pass.outcome.name), "Complete", pass.outcome.name))

ggplot(plot_data, aes(x = pass.angle, y = pass.length, color = outcome)) +
  geom_point(alpha = 0.6, size = 1.5) +  # Alpha helps see overlapping data points
  scale_x_continuous(
    breaks = c(-pi, -pi/2, 0, pi/2, pi),
    labels = c("-π", "-π/2", "0", "π/2", "π"),
    limits = c(-3.2, 3.2)
  ) +
  theme_minimal() +
  labs(
    title = "Pass Length vs. Pass Angle",
    x = "Pass Angle (radians)",
    y = "Pass Length (metres)",
    color = "Pass Outcome" # Fixes the legend title
  ) +
  theme(
    plot.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold")
  )
```

### Using complete and incomplete as labels

```{r}
plot_data <- pass_single %>%
  mutate(outcome = ifelse(is.na(pass.outcome.name), "Complete", "Incomplete"))

ggplot(plot_data, aes(x = pass.angle, y = pass.length, color = outcome)) +
  geom_point(alpha = 0.6, size = 1.5) + 
  scale_x_continuous(
    breaks = c(-pi, -pi/2, 0, pi/2, pi),
    labels = c("-π", "-π/2", "0", "π/2", "π"),
    limits = c(-3.2, 3.2)
  ) +
  theme_minimal() +
  labs(
    title = "Pass Length vs. Pass Angle",
    x = "Pass Angle (radians)",
    y = "Pass Length (yards)",
    color = "Pass Outcome"
  ) +
  theme(
    plot.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold")
  )
```

## Pass Completion compared to position on the field

```{r}
plot_data <- pass_single %>%
  mutate(outcome = ifelse(is.na(pass.outcome.name), "Complete", "Incomplete"))

ggplot(plot_data, aes(x = players_x, y = players_y, color = outcome)) +
  annotate_pitch(
    dimensions = pitch_statsbomb,
    limits = TRUE,
    alpha = 1
  ) +
  geom_point(alpha = 0.6, size = 0.8) + 
  theme_minimal() +
  labs(
    title = "Pass outcome vs position on the pitch",
    color = "Pass Outcome"
  ) +
  theme_pitch()
```

### World Cup 2022

```{r}
plot_data <- wc_2022 %>%
  mutate(outcome = ifelse(is.na(pass.outcome.name), "Complete", "Incomplete"))

ggplot(plot_data, aes(x = players_x, y = players_y, color = outcome)) +
  annotate_pitch(
    dimensions = pitch_statsbomb,
    limits = TRUE,
    alpha = 1
  ) +
  geom_point(alpha = 0.6, size = 0.8) + 
  theme_minimal() +
  labs(
    title = "Pass outcome vs position on the pitch (Men's World Cup 2022)",
    color = "Pass Outcome"
  ) +
  theme_pitch()

```

# Voronoi Diagram

```{r}
library(ggvoronoi)
library(groupdata2)
library(dplyr)
```

# Pass Sonar

-   Way of visualising pass direction and distance

-   Polar histogram

-   Length of bars indicate normalized pass frequency for each angle bin, color represents distance from each segment

-   Code credited to

```{r}
#| fig-width: 8
#| fig-height: 8
#| fig-asp: 1
#| out-width: "80%"
devtools::install_github("abhiamishra/ggshakeR")

library(ggshakeR)

bund_2324_cleaned <- bund_2324 |> 
  rename(
    x = ball_x,
    y = ball_y,
    finalX = pass_end_x,
    finalY = pass_end_y
  ) |> filter(position_group == "Goalkeeper")


plot_sonar(bund_2324_cleaned, data_type = "statsbomb", title = "Pass Sonar for Goalkeeprs")


```

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)

# 1. Prepare and bin pass data
sonar_data <- bund_2324_cleaned %>%
  mutate(
    # Calculate pass angle in radians/degrees
    angle_rad = atan2(finalY - y, finalX - x),
    angle_deg = (angle_rad * 180 / pi) %% 360,
    
    # Bin angles into 15-degree directional slices
    angle_bin = cut(angle_deg, breaks = seq(0, 360, by = 15), include.lowest = TRUE)
  ) %>%
  group_by(angle_bin) %>%
  summarise(
    total_passes = n(),
    successful_passes = sum(pass.outcome.name == "Complete", na.rm = TRUE),
    unsuccessful_passes = sum(pass.outcome.name != "Complete", na.rm = TRUE),
    avg_distance = mean(pass.length[pass.outcome.name == "Complete"], na.rm = TRUE),
    .groups = "drop"
  )

# 2. Reshape into stacked format for successful vs. unsuccessful passes
sonar_stacked <- sonar_data %>%
  pivot_longer(
    cols = c(successful_passes, unsuccessful_passes),
    names_to = "pass_type",
    values_to = "count"
  )

# 3. Plot using built-in ggplot2 continuous viridis scale
ggplot(sonar_stacked, aes(x = angle_bin, y = count, fill = avg_distance)) +
  geom_col(position = "stack", color = "#222222", linewidth = 0.2) +
  # Native ggplot2 function — no separate package required
  scale_fill_viridis_c(
    option = "magma",
    name = "Average distance (yards)",
    na.value = "#555555" # Grey for unsuccessful pass segments
  ) +
  coord_polar(start = 0) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.grid.major = element_line(color = "grey80", linetype = "dashed"),
    plot.background = element_rect(fill = "#F9F8F3", color = NA),
    panel.background = element_rect(fill = "#F9F8F3", color = NA)
  ) +
  labs(
    title = "Pass Sonar",
    x = NULL,
    y = NULL
  )
```

```{r}
library(tidyverse)
library(viridis)

round.angle=15

#Round angle to nearest 15 degrees 
bund_2324 = bund_2324 %>% mutate(angle.round=round(pass.angle*180/pi/round.angle)*round.angle)

#create data frame for sonars per player and team for a season
#this normalizes the most frequent pass angle to 1 and all other angles relative to that angle
#it also associates pass distance with each angle, but this can be changed to anything
#such as pass success or any other metric
library(tidyverse)

sonar <- bund_2324 %>%
  group_by(player.id, team.name) %>%
  mutate(N = n()) %>%
  group_by(angle.round, .add = TRUE) %>%
  mutate(n.angle = n() / N) %>%
  group_by(player.id, team.name) %>%
  mutate(
    maxN = max(n.angle),
    angle.norm = n.angle / maxN
  ) %>%
  group_by(player.id, team.name, N, angle.round) %>%
  summarize(
    angle.norm = mean(angle.norm),
    distance = mean(pass.length, na.rm = TRUE),
    distance = pmin(distance, 30),
    .groups = "drop"
  )


# Get unique player IDs
players <- unique(sonar$player.id)

# Select player index (e.g., 57th player in your list)
idx <- 57
target_id <- players[idx]

# Plot
ggplot(sonar %>% filter(player.id == target_id)) +
  geom_bar(
    aes(x = angle.round, y = angle.norm, fill = distance), 
    stat = "identity",
    width = 15 # Adjust to match your angle bin width
  ) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(
    breaks = seq(-180, 180, by = 90), 
    limits = c(-180, 180)
  ) +
  coord_polar(start = pi, direction = 1) +
  scale_fill_viridis_c("Distance (yards)", limits = c(0, 30), na.value = "#FDE725FF") +
  labs(
    x = NULL, 
    y = NULL, 
    title = paste("Player ID:", target_id)
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 12),
    plot.background = element_rect(fill = "transparent", colour = NA),
    panel.background = element_rect(fill = "transparent", colour = NA)
  )

#Plotting on top of a field by a team's formation
#The trick is to save each players PassSonar as a grob into a list. Then using
#annotation_custom() each PassSonar is placed in the correct position on the pitch
#It takes some trial and error to get the PassSonars into the correct position

text_color="black"
background_color="white"
radar.size=27
ymax=80
xmax=120

team.select="North Carolina Courage"
match.select=7444

game.lineup=StatsBombData%>%filter(team.name==team.select, type.name=='Starting XI', match_id==match.select )
game.players=game.lineup$tactics.lineup[[1]][["player.name"]]
team.formation= parse_number( game.lineup$tactics.formation)
#game.lineup$tactics.lineup[[1]][["position.name"]]  #uncomment to view positions to help place into correct locations of field

player.plots=list()
for (i in 1:length(game.players)){
  
  plot.data=sonar%>%filter(team.name==team.select & player.name==game.players[i])
  
  player.plots[[i]]=ggplot(plot.data)+geom_bar(aes(x=angle.round, y=angle.norm, fill=distance), stat="identity")+
    scale_y_continuous(limits=c(0,1))+
    scale_x_continuous(breaks=seq(-180,180, by=90), limits=c(-180,180))+
    coord_polar(start=pi, direction=1)+
    scale_fill_viridis("Distance (yards)", limits=c(0,30), na.value="#FDE725FF")+
    labs(x='', y='',title=plot.data$player.name[1])+
    theme_void()+
    theme(plot.title = element_text(hjust=0.5, color=text_color),
          plot.background = element_rect(fill = "transparent",colour = NA),
          panel.background = element_rect(fill = "transparent",colour = NA),
          legend.position = "none")
  player.plots[[i]]=ggplotGrob(player.plots[[i]])
  
  if (i==length(game.players)){
    colorbar=
      ggplot(plot.data)+geom_bar(aes(x=angle.round, y=angle.norm, fill=distance), stat="identity")+
      scale_y_continuous(limits=c(0,0))+
      scale_fill_viridis("", limits=c(0,30), na.value="#FDE725FF")+
      labs(x='', y='')+
      theme_void()+
      theme( legend.position = "bottom",
             plot.background = element_rect(fill = "transparent",colour = NA),
             panel.background = element_rect(fill = "transparent",colour = NA))
    colorbar=ggplotGrob(colorbar)
  }
}

#this is a 4-4-2 example. Use similar methods for other formations

if (team.formation==442){
  team.formation='4-4-2'
  
  back.line=20
  mid.line=48
  forward.line=77
  p=createPitch(grass_colour = background_color,goal_colour=text_colour, line_colour = text.color)+coord_flip(ylim=c(0,80))+
    theme(aspect.ratio = 120/80, plot.title = element_text(size=18, hjust=0.5, vjust=-2, color=text.color),
          plot.background = element_rect(fill = background_color,colour = NA),
          panel.background = element_rect(fill = background_color,colour = NA))+
    annotation_custom(grob=player.plots[[1]], xmin=-9, xmax=-9+radar.size, ymax=ymax/2+radar.size/2-1.5, y=ymax/2-radar.size/2-1.5)+ #GK
    annotation_custom(grob=player.plots[[2]], xmin=back.line+3, xmax=back.line+3+radar.size, ymax=ymax+1, y=ymax-radar.size+1)+ #RB
    annotation_custom(grob=player.plots[[5]], xmin=back.line+3, xmax=back.line+3+radar.size, ymax=-3+radar.size, y=-3)+ #LB
    annotation_custom(grob=player.plots[[4]], xmin=back.line, xmax=back.line+radar.size, ymax=ymax/2-23.5+radar.size, y=ymax/2-23.5)+ #LCB
    annotation_custom(grob=player.plots[[3]], xmin=back.line, xmax=back.line+radar.size, ymax=ymax/2-6+radar.size, y=ymax/2-6)+ #RCB
    annotation_custom(grob=player.plots[[7]], xmin=mid.line, xmax=mid.line+radar.size, ymax=ymax/2-23.5+radar.size, y=ymax/2-23.5)+ #LCM
    annotation_custom(grob=player.plots[[6]], xmin=mid.line, xmax=mid.line+radar.size, ymax=ymax/2-6+radar.size, y=ymax/2-6)+ #RCM
    annotation_custom(grob=player.plots[[10]], xmin=forward.line, xmax=forward.line+radar.size, ymax=ymax/2-3+radar.size, y=ymax/2-3)+ #RF
    annotation_custom(grob=player.plots[[11]], xmin=forward.line, xmax=forward.line+radar.size, ymax=ymax/2-26+radar.size, y=ymax/2-26)+ #LF
    annotation_custom(grob=player.plots[[8]], xmin=mid.line+5, xmax=mid.line+5+radar.size, ymax=ymax+1, y=ymax-radar.size+1)+ #RM
    annotation_custom(grob=player.plots[[9]], xmin=mid.line+5, xmax=mid.line+5+radar.size, ymax=-3+radar.size, y=-3)+ #LM
    annotation_custom(grob=colorbar, xmin=3, xmax=7, ymin=1, ymax=18)+
    annotate("text", label="concept:@etmckinley\ndata:@StatsBomb", x=6, y=79, hjust=1,vjust=1 ,size=3.75, color=text.color)+
    annotate("text", label="Mean Pass Distance (Yards)", x=9, y=3, hjust=0, size=3, color=text.color)+
    annotate("text", label='Bar length = normalized pass angle frequency; Bar color = mean pass distance', color=text.color, x=-2, y=79, hjust=1, size=3)+
    annotate("text", label=paste0('Starting Formation: ', team.formation), color=text.color, x=-2, y=0, hjust=0, size=5, fontface="bold")+
    annotate("text", label=paste0('PassSonar: ', team.select), color=text.color, x=121.5, y=0, hjust=0, size=9, fontface="bold")+
    guides(fill = guide_colourbar())
  
}

ggsave(p, file=paste0('./', team.select,' PassSonar.png'), width=9.5, height=11.5, dpi=150, bg=background_color )


```

### Voronoi Diagrams

```{r}
# #| label: Voronoi Diagrams
# library(deldir)
# library(sf)
# 
# testframe = head(open_play_passes,16) # Test frame
# 
# library(deldir)
# library(sf)
# library(purrr)
# 
# get_visible_voronoi_areas <- function(player_positions, visible_area_vec) {
#   va_coords <- matrix(visible_area_vec, ncol = 2, byrow = TRUE)
# 
#   vor <- deldir(player_positions$x, player_positions$y, rw = c(0, 120, 0, 80))
#   tiles <- tile.list(vor)
# 
#   cell_polys <- lapply(tiles, function(t) {
#     coords <- cbind(t$x, t$y)
#     coords <- rbind(coords, coords[1, ])
#     st_polygon(list(coords))
#   })
#   cell_sf <- st_sf(player_id = seq_along(cell_polys),
#                     geometry = st_sfc(cell_polys))
# 
#   va_closed <- rbind(va_coords, va_coords[1, ])
#   visible_poly <- st_sfc(st_polygon(list(va_closed)))
# 
#   clipped <- st_intersection(cell_sf, st_sf(geometry = visible_poly))
#   clipped$area <- st_area(clipped)
#   clipped
# }
# 
# result <- get_visible_voronoi_areas(
#   player_positions = data.frame(x = testframe$players_x, y = testframe$players_y),
#   visible_area_vec = testframe$visible_area[[1]]   # any row works, they're all identical
# )
# 
# result
# 
# library(ggplot2)
# library(ggsoccer)
# 
# # add team/player info back for coloring
# result$teammate <- testframe$teammate
# result$actor <- testframe$actor
# 
# ggplot() +
#   annotate_pitch(dimensions = pitch_statsbomb, colour = "white", fill = "#1c6b3c") +
#   geom_sf(data = result, aes(fill = teammate), colour = "black", alpha = 0.5, inherit.aes = FALSE) +
#   geom_point(data = testframe, 
#              aes(x = players_x, y = players_y, colour = teammate), 
#              size = 4) +
#   geom_point(data = subset(testframe, actor == TRUE),
#              aes(x = players_x, y = players_y),
#              shape = 21, size = 7, stroke = 1.5, colour = "gold", fill = NA) + 
#   scale_fill_manual(values = c("TRUE" = "blue", "FALSE" = "red")) +
#   scale_colour_manual(values = c("TRUE" = "darkblue", "FALSE" = "darkred")) +
#   theme_pitch() +
#   coord_sf() +
#   labs(title = "Voronoi cells clipped to visible area")
```

```{r}

```

# Appendix

## Full Data Description

## Number of passes by position, season and competion

```{r}

```
